\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-06 paper.tex
Rewrite the 'Clean the table' study

\begin{tabular}{llll}
 \bf{Study} & Focus & \emph{Notes} & Reference \\
\hline
{\it Point \& Learn} (2010) & Interactive knowledge acquisition & & \cite{Lemaignan2010} \\
{\it Spy Game} (2010) & Interactive object discrimination & & \cite{Ros2010b} \\
{\it Interactive Grounding I} (2011) & Multi-modal interaction, perspective taking & \emph{immobile robot} & \cite{lemaignan2011what} \\
{\it Roboscopie} (2011) & Theater, reflection on the future of HRI & \emph{semi-autonomous} & \cite{lemaignan2012roboscopie} \\
{\it Cleaning the table} (2011) & Whole stack integration & & \cite{Alami2011} \\
{\it I'm in your shoes} (2012) & False beliefs & & \cite{warnier2012when} \\
{\it Give me this} (2012) & Natural joint object manipulation & & \cite{gharbi2013natural} \\
{\it Interactive Grounding II} (2012) & Multi-modal interaction, perspective taking & & \cite{lemaignan2013talking} \\
\hline

@@

\subsection{\emph{Cleaning the table} scenario}

This experiment involved a more complex decisional layer where the {\sc Oro} server
was used in conjunction with the HATP symbolic task planner and the {\sc Shary}
execution controller (the complete software architecture we deployed for this
experiment is the one pictured in Figure~\ref{fig|archi},
page~\pageref{fig|archi}).

In this scenario (Figure~\ref{fig|cleantable-video}), a human and a robot
cooperate to remove objects from a table. The robot produces symbolic plans for
both itself and the human (Figure~\ref{plan-etat2})
that allow the robot to (verbally) share the task with the human (like ``I take
the green box and I put it in the trashbin, you take the black video tape and
you throw it in the trashbin''). Plans are created based on perceived
visibility and reachability of the objects, and the robot also monitors the
human activities to track the advancement of the whole plan.

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{cleantable.jpg}

    \caption{Snapshot of the film of the ``Clean the Table'' scenario. The
    physical situation, the {\sc Spark} model, and the current step of the plan can
    be seen.}

@@

\begin{figure*}[thpb]
  \centering
\resizebox{0.9\textwidth}{!}{%
\begin{tikzpicture}[
    >=latex,
    robot/.style={fill=RoyalBlue!50},
    every edge/.style={<-, draw, ultra thick},
    every node/.style={ draw, 
                        thick,  
                        circle, 
                        font=\sf,
                        align=center,
                        node distance=2cm,
                        fill=BurntOrange!50, 
                        minimum size=2cm, 
                        inner sep=0.1cm}]

    \node (h1) {\action{TAKE}{WALLE\_TAPE}{TABLE}};
    \node[right=of h1] (h2) {\action{THROW}{WALLE\_TAPE}{TRASH\_1 }} edge (h1);

    \node[robot, below=1.5 of h1] (r1) {\action{TAKE}{GREY\_TAPE}{TABLE }};
    \node[robot, right=of r1] (r2) {\action{PUTRV}{GREY\_TAPE}{TABLE}} edge (r1);
    \node[robot, right=of r2] (r3) {\action{TAKE}{LOTR\_TAPE}{TABLE }} edge (r2);
    \node[robot, right=of r3] (r4) {\action{PUTRV}{LOTR\_TAPE}{TABLE }} edge (r3);

    \node[right=of h2] (h3) {\action{TAKE}{GREY\_TAPE}{TABLE}} edge (h2) edge (r2);
    \node[right=of h3] (h4) {\action{THROW}{GREY\_TAPE}{TRASH\_1}} edge (h3);
    \node[right=of h4] (h5) {\action{TAKE}{LOTR\_TAPE}{TABLE}} edge (h4) edge (r4);
    \node[right=of h5] (h6) {\action{THROW}{LOTR\_TAPE}{TRASH\_1}} edge (h5);

  \caption {A plan produced by HATP to execute the high-level order ``clean the
  table''. Two streams of actions are generated: for the human (top) and for the
  robot (bottom). Synchronization points ensure the coordination.}

  \label{plan-etat2}
\end{figure*}

Figure~\ref{fig|cleantable-timeline} presents a sample of the whole task and
shows how the different components produce and use symbolic knowledge.

This sequence of events depicts a run with only one tape on the table. The tape is
reachable by the robot only, while the human is the one who can reach the
trashbin to throw objects.

@@

        \node at (2,0) (percept) {\bf Perception};
            \node[below=0.5 of percept.south west, anchor=mid] {camera};
            \node[below=0.5 of percept.south east, anchor=mid] {3D model};
        \node[right=4 of percept, minimum width=2.5cm] (kb) {\bf Knowledge};
            \node[below=0.5 of kb.south west, anchor=mid east] (kbr) {robot};
            \node[below=0.5 of kb.south east, anchor=mid west] (kbh) {human};
            \draw[dotted] (kb.south) to (figbottom -| kb);
        \node[right=4 of kb] (plan) {\bf Plans};
            \node[below=0.5 of plan.south west, anchor=mid east] (probot) {robot};
            \node[below=0.5 of plan.south east, anchor=mid west] (phuman) {human};
        \node[right=5 of plan] (action) {\bf Actions};
            \node[below=0.5 of action.south west, anchor=mid east] (arobot) {robot};
            \node[below=0.5 of action.south east, anchor=west] (ahuman) {human\\(monitoring)};

@@

        \node at (pr1 -| arobot) (ep1) {\it evaluate preconditions};

        \node[below=0.1 of ep1] (mhp1) {%
            \begin{tikzpicture}
                \node (title) {\bf motion planning};
                \node[below=0.1 of title.south west, label=below:{\tt PICK\_GOTO}] (mapg) {\includegraphics{MHP_ARM_PICK_GOTO}};
                \node[right=of mapg, label=below:{\tt TAKE\_TO\_FREE}] (mattf) {\includegraphics{MHP_ARM_TAKE_TO_FREE}} edge[<-] (mapg);
            \end{tikzpicture}
        };
        \node[below=0.1 of mhp1] (me1) {\bf motion execution};
        \node[below=0.1 of me1] (ae1) {\it assess effects};

        %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \node at (pr2 -| arobot) (ep2) {\it evaluate preconditions};

        \node[below=0.1 of ep2] (mhp2) {%
            \begin{tikzpicture}
                \node (title) {\bf motion planning};
                \node[below=0.1 of title, label=below:{\tt ESCAPE}] (maeo) {\includegraphics{MHP_ARM_ESCAPE_OBJECT}};
                \node[left=of maeo, label=below:{\tt PLACE\_FROM\_FREE}] (mapff) {\includegraphics{MHP_ARM_PLACE_FROM_FREE}} edge (maeo);
                \node[right=of maeo, label=below:{\tt TO\_FREE}] (maf) {\includegraphics{MHP_ARM_FREE}} edge[<-] (maeo);
            \end{tikzpicture}
        };
        \node[below=0.1 of mhp2] (me2) {\bf motion execution};
        \node[below=0.1 of me2] (ae2) {\it assess effects};

@@

 \end{tikzpicture}
 }
 \caption {Sequence of events of a simplified version of the \emph{Cleaning the table} scenario.}
  \label{fig|cleantable-timeline}
\end{figure*}

The plan produced in that case by HATP is straightforward and shown in the
third row. It consists of four successive actions involving the robot and the
human. Robot grasps the tape and then places it on the table at a position
where it is visible and reachable for the human. The human is then asked to pick
the tape and throw it in the trashbin.

>>>>

\begin{tabular}{lll}
 \bf{Study} & Focus & Reference \\
\hline
{\it Point \& Learn} (2010) & Interactive knowledge acquisition & \cite{Lemaignan2010} \\
{\it Spy Game} (2010) & Interactive object discrimination & \cite{Ros2010b} \\
{\it Interactive Grounding I} (2011) & Multi-modal interaction, perspective taking & \cite{lemaignan2011what} \\
{\it Roboscopie} (2011) & Theater, reflection on the future of HRI & \cite{lemaignan2012roboscopie} \\
{\it Cleaning the table} (2011) & Whole stack integration & \cite{Alami2011} \\
{\it I'm in your shoes} (2012) & False beliefs & \cite{warnier2012when} \\
{\it Give me this} (2012) & Natural joint object manipulation & \cite{gharbi2013natural} \\
{\it Interactive Grounding II} (2012) & Multi-modal interaction, perspective taking & \cite{lemaignan2013talking} \\
\hline

@@

\subsection{\emph{Cleaning the table} study}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.6\columnwidth]{cleantable.jpg}

    \caption{Face-to-face setting of the \emph{Clean the Table} scenario. The
    physical situation, the {\sc Spark} model, and the current step of the plan
    are visible on the picture.}

@@

This study demonstrates a richer decision-making process. The {\sc Oro} server
is used in conjunction with the HATP symbolic task planner and the {\sc Shary}
execution controller to produce and execute a shared plan.

%In this scenario (Figure~\ref{fig|cleantable-video}), a human and a robot
%cooperate to remove objects from a table. The robot produces symbolic plans for
%both itself and the human (Figure~\ref{plan-etat2})
%that allow the robot to (verbally) share the task with the human (like ``I take
%the green box and I put it in the trashbin, you take the black video tape and
%you throw it in the trashbin''). Plans are created based on perceived
%visibility and reachability of the objects, and the robot also monitors the
%human activities to track the advancement of the whole plan.

%
%\begin{figure*}[thpb]
%  \centering
%\resizebox{0.9\textwidth}{!}{%
%\begin{tikzpicture}[
%    >=latex,
%    robot/.style={fill=RoyalBlue!50},
%    every edge/.style={<-, draw, ultra thick},
%    every node/.style={ draw, 
%                        thick,  
%                        circle, 
%                        font=\sf,
%                        align=center,
%                        node distance=2cm,
%                        fill=BurntOrange!50, 
%                        minimum size=2cm, 
%                        inner sep=0.1cm}]
%
%    \node[font=\rm\large, draw=none, fill=none, rotate=90] at (-2, 0) {Human\\actions};
%    \node[font=\rm\large, draw=none, fill=none, rotate=90] at (-2, -3.5) {Robot\\actions};
%
%
%    \node (h1) {\action{TAKE}{WALLE\_TAPE}{TABLE}};
%    \node[right=of h1] (h2) {\action{THROW}{WALLE\_TAPE}{TRASH\_1 }} edge (h1);
%
%    \node[robot, below=1.5 of h1] (r1) {\action{TAKE}{GREY\_TAPE}{TABLE }};
%    \node[robot, right=of r1] (r2) {\action{PUTRV}{GREY\_TAPE}{TABLE}} edge (r1);
%    \node[robot, right=of r2] (r3) {\action{TAKE}{LOTR\_TAPE}{TABLE }} edge (r2);
%    \node[robot, right=of r3] (r4) {\action{PUTRV}{LOTR\_TAPE}{TABLE }} edge (r3);
%
%    \node[right=of h2] (h3) {\action{TAKE}{GREY\_TAPE}{TABLE}} edge (h2) edge (r2);
%    \node[right=of h3] (h4) {\action{THROW}{GREY\_TAPE}{TRASH\_1}} edge (h3);
%    \node[right=of h4] (h5) {\action{TAKE}{LOTR\_TAPE}{TABLE}} edge (h4) edge (r4);
%    \node[right=of h5] (h6) {\action{THROW}{LOTR\_TAPE}{TRASH\_1}} edge (h5);
%
%
%\end{tikzpicture}
%}
%
%  \caption {A plan produced by HATP to execute the high-level order ``clean the
%  table''. Two streams of actions are generated: for the human (top) and for the
%  robot (bottom). Synchronization points ensure the coordination.}
%
%  \label{plan-etat2}
%\end{figure*}
%
Figure~\ref{fig|cleantable-timeline} presents a practical sample of the whole
task. It depicts a run with a single tape on a table. The tape is
reachable by the robot only, while the bin where objects are supposed to be
placed is reachable by the human only: the robot needs to come up with a shared
plan that involves joint actions.

The goal is first received by the execution controller (after processing of the
user request by the {\sc Dialogs} module, not shown on the figure). At
$t_1$, the tape is computed by the robot as being reachable by the robot only
(columns \emph{Perception} and \emph{Knowledge}), and the execution controller
invokes the task planner, which produces a joint plan (column \emph{Plan}).

The first task ({\tt TAKE(GREY\_TAPE, TABLE)}) is instantiated by calling the 3D
motion planner (column \emph{Actions}, left). The motion planner returns two atomic
actions ({\tt PICK\_GOTO} followed by {\tt TAKE\_TO\_FREE}) that the controller
executes (by first reaching the object, grasping it and bringing it back to a
free position).  Once the robot's perception allows the controller to ensure
that the task's post-conditions are fulfilled (at $t_2$, by satisfying the
statement \stmt{ROBOT hasInHand TAPE}), the next task is started (placing the
tape so that it becomes reachable by the human).

At $t_3$, the tape is now reachable by the human, and the next tasks (taking the
tape and placing it in the bin) have to be performed by the human: the robot
instructs the user to do so (not pictured on the figure) and monitors the
actions of the human to detect when the tasks' post-conditions are satisfied
(column \emph{Actions}, right). When these post-conditions are fulfilled, the
goal is considered to be completed.

This simple example illustrates how the symbolic facts are produced from the
situation assessment, and, in parallel, used by the execution controller to
assess the overall progress of the plan.

@@

        \node at (2,0) (percept) {\bf Perception};
            \node[below=0.5 of percept.south west, anchor=mid] {camera};
            \node[below=0.5 of percept.south east, anchor=mid] {3D model};
        \node[right=4 of percept, minimum width=2.5cm] (kb) {\bf Knowledge};
            \node[below=0.5 of kb.south west, anchor=mid east] (kbr) {robot};
            \node[below=0.5 of kb.south east, anchor=mid west] (kbh) {human};
            \draw[dotted] (kb.south) to (figbottom -| kb);
        \node[right=4 of kb] (plan) {\bf Plan};
            \node[below=0.5 of plan.south west, anchor=mid east] (probot) {robot};
            \node[below=0.5 of plan.south east, anchor=mid west] (phuman) {human};
        \node[right=5 of plan] (action) {\bf Actions};
            \node[below=0.5 of action.south west, anchor=mid east] (arobot) {robot};
            \node[below=0.5 of action.south east, anchor=west] (ahuman) {human\\(monitoring)};

@@

        \node at (pr1 -| arobot) (ep1) {\it evaluate pre-conditions};

        \node[below=0.1 of ep1] (mhp1) {%
            \begin{tikzpicture}
                \node (title) {\bf motion planning};
                \node[below=0.1 of title.south west, label=below:{\tt PICK\_GOTO}] (mapg) {\includegraphics{MHP_ARM_PICK_GOTO}};
                \node[right=of mapg, label=below:{\tt TAKE\_TO\_FREE}] (mattf) {\includegraphics{MHP_ARM_TAKE_TO_FREE}} edge[<-] (mapg);
            \end{tikzpicture}
        };
        \node[below=0.1 of mhp1] (me1) {\bf motion execution};
        \node[below=0.1 of me1] (ae1) {\it assess post-conditions};

        %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \node at (pr2 -| arobot) (ep2) {\it evaluate pre-conditions};

        \node[below=0.1 of ep2] (mhp2) {%
            \begin{tikzpicture}
                \node (title) {\bf motion planning};
                \node[below=0.1 of title, label=below:{\tt ESCAPE}] (maeo) {\includegraphics{MHP_ARM_ESCAPE_OBJECT}};
                \node[left=of maeo, label=below:{\tt PLACE\_FROM\_FREE}] (mapff) {\includegraphics{MHP_ARM_PLACE_FROM_FREE}} edge (maeo);
                \node[right=of maeo, label=below:{\tt TO\_FREE}] (maf) {\includegraphics{MHP_ARM_FREE}} edge[<-] (maeo);
            \end{tikzpicture}
        };
        \node[below=0.1 of mhp2] (me2) {\bf motion execution};
        \node[below=0.1 of me2] (ae2) {\it assess post-conditions};

@@

 \end{tikzpicture}
 }
 \caption {Chronology of the \emph{Cleaning the table} scenario. The light
     arrows indicate the general flow. This figure
     represents a simplified version: only one object, {\tt TAPE}, must
     be removed from the table and placed in the bin {\tt BIN}.  }
  \label{fig|cleantable-timeline}
\end{figure*}

%The plan produced in that case by HATP is straightforward and shown in the
%third row. It consists of four successive actions involving the robot and the
%human. Robot grasps the tape and then places it on the table at a position
%where it is visible and reachable for the human. The human is then asked to pick
%the tape and throw it in the trashbin.
